Trygve's Problem

### 2018-03-20 generativemodelsforscrna/outline.tex
updates for mtg

\begin{description}
  \item[Status] Succesfully trained the new ``iterative model,'' evaluated it to some extent
  \item[Todo] Compare clusters against Allen, try unamortized inference
\end{description}

@@

Here HNBPF is a negative-binomial-poisson-factorization approach, ZINB-Vae is based on a Michael Jordan paper, $k$ is the number of latent dimensions.  The results are certainly not mind-blowing, but its nice to see that we're at least heading in a reasonable direction.  There may be an upper limit on the PMP, due to very real technical noise in the system, so it is hard to evaluate whether this is a trivial improvement or not.

To arrive at our figures for our model, we use a truly ridiculous method of variational inference.  In particular, we simply \emph{zero out} all of the ``unobserved'' entries, and apply our inference network unchanged to get estimates for the variational parameters.  This is not ``cheating'' in the sense that it is a practical method for imputation in the case where certain genes are unobserved.  However, it might well do less than optimally.  I was talking to Scott, and it sounds like the question of amortized inference with unpredictably missing data is a big deal right now.  Worth thinking about more.

\subsection{Denoising}

>>>>

\begin{description}
  \item[Status] Succesfully trained the new ``iterative model,'' evaluated it to some extent
  \item[Todo] Compare clusters against Allen, consider linear model, do more layers
\end{description}

@@

Here HNBPF is a negative-binomial-poisson-factorization approach, ZINB-Vae is based on a Michael Jordan paper, $k$ is the number of latent dimensions.  The results are certainly not mind-blowing, but its nice to see that we're at least not doing worse.

To do posterior inference on held out likelihood, our approach is to first set all of the ``unobserved'' entries to the average value for the corresponding genes, and apply our inference network unchanged to get estimates for the variational parameters.  We then refine these estimates with gradient descent.  We note that if we use completely random guesses to initialize the gradient descent, we often converge to inferior local minima; this shows the benefit of the amortized inference.

\subsection{Posterior predictive samples}

One limitation of the method as we have implemented it is that the posterior predictive distribution seems to be substantially underfitting.  To see what we mean, consider the following figure:

\includegraphics[width=0.8\textwidth]{pics/Z1gn}

Here we four rows and three columns of plots.  Each row corresponds to a different gene.  Each dot in each plot corresponds to a cell.  The position of the dot indicates the latent representation of the cell, as inferred by posterior inference.  The color indicates the amount of expression for the particular gene for that cell, measured three ways:

\begin{enumerate}
\item The first column shows the true expression of the appropriate gene for each cell (i.e the color of each dot is the gene expression for that cell)
\item The second column shows a posterior predictive sample of the gene expression for each cell
\item The third column shows the mean of such samples for each cell
\end{enumerate}

We see that the true expression doesn't look much like the posterior predictive expression for the Vip and Sst genes.  Something better is needed.  Possibly more layers.  Possibly more training.  Worth thinking about.

### 2018-03-20 trygve/outline.tex
generatiev models
updates for mtg

@@

...look reasonable, and more or less checked out with the Allen Institute people.  That said, they caution that some cell types tend to die more often in one experimental modality than another, which would violate our assumptions.  However, this doesn't bother them, because they see this technique as a way to \emph{validate} their assignments of clusters.  As such, if the result of our algorithm doesn't make sense, that's great because it may help them to figure out \emph{which cell types are dying more} and then they can upweight certain groups correspondingly.  Or it may help them figure out that their cluster assignment was all wrong.  This is how they see my method fitting into their pipelines: as a way of sanity checking their cluster assignments.  Which seems cool.

>>>>

\begin{description}
  \item[Status] Designed estimator, showed result to Allen people, they seem happy
  \item[Todo] Prove consistency, try on other Allen datasets, sample uniformly from possible set
\end{description}

@@

We looked at Allen Institute cell-type assignment of cells, performed using two different modalities: facs and patchseq.  Our input was two table: cre lines by facs cluster, and cre lines by patch cluster (Figure \ref{fig:alleninput}).  Our output was facs cluster by patch cluster (Figure \ref{fig:allenoutput}).

\begin{figure}
\includegraphics[width=.9\textheight,angle=270]{pics/alleninput}
\caption{Allen institute data: Our input \label{fig:alleninput}}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{pics/allen}
\caption{Allen institute data: Our output \label{fig:allenoutput}}
\end{figure}

These looked reasonable to the scientists, and also pointed out some possible issues.  In particular, some cell types tend to die more often in one experimental modality than another, which would violate our assumptions.  However, this doesn't bother them, because they see this technique as a way to \emph{validate} their assignments of clusters.  As such, if the result of our algorithm doesn't make sense, that's great because it may help them to figure out \emph{which cell types are dying more} and then they can upweight certain groups correspondingly.  Or it may help them figure out that their cluster assignment was all wrong.  This is how they see my method fitting into their pipelines: as a way of sanity checking their cluster assignments.  Which seems cool.

There is one possibly confusing aspect of our output that might be nice to address.  The upper and lower bounds (as we have visualized them) are each actually impossible as values for $q$.  For example, the rows of the upper bounds generally sum to more than 1.  This is because we have taken \emph{elementwise bounds} on each entry.  However, it would be nice to sample uniformly from the set of possible values of $q$ which are consistent with our observations.  MCMC or rejection sampling might be good options to investigate.  These would help give the user another way to understand what possible values $q$ might take on.
